### 2017-05-01 article.tex
More text in the article.

\documentclass{article}

\pdfminorversion=4

\usepackage{tikz}
\usetikzlibrary{fit,patterns,decorations.pathreplacing}

\usepackage{graphicx}

\def\thindivider{\centerline{\tiny%%
-- --- --- -- ~~~ --- --- --- ~~~ --- -- --- -- ~~~~~~
--- --- --- ~~~  -- --- -- ~~~~~~
--- --- -- ~~~ --- ~~~ -- -- --- -- ~~~ --- --- ---}}

\begin{document}

\title{This PDF is a Git Repository\\Containing its Own \LaTeX\ Source}
\author{Evan Sultanik}
\date{April 11, 2017}

\maketitle

Have you ever heard of the \texttt{git bundle} command? I hadn't. It
bundles a set of Git objects---potentially even an entire
repository---into a single file. Git allows you to treat that file as
if it were a standard Git database, so you can do things like clone a
repo directly from it. Its purpose is to easily sneakernet pushes or
even whole repositories across air gaps.

\thindivider

Neighbors, it's possible to create a PDF that is also a Git
repository.

\begin{center}
\begingroup
\setbox9=\hbox{\footnotesize\verb|Receiving objects: 100% (174/174), 103.48 KiB | 0 bytes/s, done.|}
\begin{minipage}{\wd9}
\footnotesize\begin{verbatim}
$ git clone article.pdf foo
Cloning into 'foo'...
Receiving objects: 100% (174/174), 103.48 KiB | 0 bytes/s, done.
Resolving deltas: 100% (100/100), done.
$ cd foo
$ ls
article.tex
article.pdf
\end{verbatim}
\end{minipage}
\endgroup
\end{center}

\section{The Git Bundle File Format}

The file format for Git bundles doesn't appear to be formally
specified anywhere, however, inspecting \texttt{bundle.c} reveals that
it's relatively straightforward:
\def\returnkey{\hspace*{1pt}{\setbox9=\hbox{$\hookleftarrow$}\tikz\node[text width=\wd9,fill=lightgray,rounded corners=1pt] at (0,0){$\hookleftarrow$};}\hspace*{1pt}}
\begin{center}
\begin{tikzpicture}
  \node[text width=0.9\hsize] at (0,0) (sig) {\footnotesize\tt\# v2 git bundle{\tiny\returnkey}};
  \node[inner sep=0mm,fit=(sig),draw,rounded corners,dashed,minimum width=0.9\hsize] (sigbox) {};
  \node[draw,rounded corners,fill=white] at (sigbox.north) (siglabel) {\footnotesize Git Bundle Signature};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (sig.south west) (b1) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/heads/master{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b1.south west) (b2) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/remotes/origin/master{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b2.south west) (b3) {\footnotesize\tt 4146cfe2fe9249fc14623f832587efe197ef5d2d refs/stash{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b3.south west) (b4) {\footnotesize\tt babdda4735ef164b7023be3545860d8b0bae250a HEAD{\tiny\returnkey}};
  \node[inner sep=0mm,fit=(b1) (b4),draw,rounded corners,dashed,minimum width=0.9\hsize] (digestbox) {};
  \node[rotate=-90,draw,rounded corners,fill=white] at (digestbox.east) {\footnotesize Digest};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b4.south west) (empty) {{\tiny\returnkey}\footnotesize\ };
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (empty.south west) (pack) {{\footnotesize\tt PACK}$\ldots$};
  \node[inner sep=0mm,fit=(pack),draw,rounded corners,dashed,minimum width=0.9\hsize] (packbox) {};
  \node[draw,rounded corners,fill=white] at (pack.south) {\footnotesize Git Packfile};
\end{tikzpicture}
\end{center}
Git uses its Packfile format to compress the objects in its database,
as well as to reduce network bandwidth when pushing and pulling, which
makes it an obvious choice for use inside bundles. This of course
raises the question: What is the format for a Git Packfile?

Git does have some internal documentation in
\begin{center}
  \texttt{Documentation/technical/pack-format.txt}
\end{center}
however, it is rather sparse, and does not provide enough detail to
fully parse the format. The documentation also has some
``observations'' that suggest it wasn't even written by the file
format's creator and instead was written by a developer who was later
trying to make sense of the code.

Luckily, Aditya Mukerjee already had to reverse engineer the file
format for his GitGo clean-room implementation of Git, and he wrote an
excellent blog entry about
it\footnote{\texttt{https://codewords.recurse.com/issues/three/unpacking-git-packfiles}}.

\begin{center}
{\small
\def\char#1{\,`\texttt{#1}'\,}
\setbox9=\hbox{\char{K}}
\newdimen\bytewidth\bytewidth=\wd9
\def\byte#1{\hbox to \bytewidth{\hfil\texttt{#1}\hfil}}
\def\desc#1#2{\hbox to #1\bytewidth{\hfil #2\hfil}}
\def\underbrace#1{\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=2pt
    },
    decorate
] ([xshift=1pt]#1.base west) -- ([xshift=-1pt]#1.base east) 
node (#1label) [pos=0.5,anchor=north,yshift=-2pt]}
\begin{tikzpicture}
\node[inner sep=0pt,minimum width=4\bytewidth] at (0,0) (pack) {\char{P}\char{A}\char{C}\char{K}};
\node[inner sep=0pt,anchor=base west] at (pack.base east) (version) {\byte{00}\byte{00}\byte{00}\byte{02}};
\node[inner sep=0pt,anchor=base west] at (version.base east) (numobj) {\desc{4}{\# objects}};
\underbrace{pack}{\tiny magic};
\underbrace{version}{\tiny version};
\underbrace{numobj}{\tiny big-endian 4 byte int};
\node[inner sep=0pt,anchor=north west] at (packlabel.south -| pack.west) (chunks) {one data chunk for each object};
\node[inner sep=0pt,anchor=north west] at ([yshift=-0.5\baselineskip]chunks.south west) (sha1) {20-byte SHA-1 of all the previous data in the pack};
\end{tikzpicture}}
\end{center}

I hope you like Soviet holes, dear neighbor, because chasing this
rabbit might remind you of \includegraphics{kolskaya}. If you
grew bored reading Tristram Shandy, it's safe to skip to the next
section, 'cause it's turtles all the way down.

\begin{center}
\includegraphics[width=0.5\hsize]{RazvodityeKrolikov_small}
\end{center}

Okay, the next step is to figure out the ``chunk'' format.  The chunk
header is variable length, and can be as small as one byte. It encodes
the type of object and the \emph{uncompressed} size of the object. If
the object is a \textit{delta} (\textit{i.e.}, a diff, as opposed to a
complete object), the header is followed by either the SHA-1 hash of
the base object to which the delta should be applied, or a byte
reference within the packfile for the start of the base object. The
remainder of the chunk consists of the object data, zlib-compressed.

This is the format of the variable length chunk header:
\begin{center}
{\newcount\bitnum\bitnum=0
\begin{tikzpicture}
\node[coordinate] at (0,0) (b0) {};
\foreach \i in {1,0,1,1,0,1,0,0,1,1,0,1,0,1,1,0,0,1,0,0,1,0,0,1} {
  {\newcount\prevbit\prevbit=\bitnum
  \global\advance\bitnum by 1
  \node[anchor=west] at (b\the\prevbit.east) (b\the\bitnum) {\i};}
  \draw ([xshift=1pt]b\the\bitnum.south west) -- ([xshift=-1pt]b\the\bitnum.south east);
}
\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=-2pt
    },
    decorate
] ([xshift=-1pt]b8.north east) -- ([xshift=1pt]b1.north west)
node [pos=0.5,anchor=south,yshift=-2pt] {\footnotesize first byte};
\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=-2pt
    },
    decorate
] ([xshift=-1pt]b16.north east) -- ([xshift=1pt]b9.north west)
node [pos=0.5,anchor=south,yshift=-2pt] {\footnotesize second byte};
\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=-2pt
    },
    decorate
] ([xshift=-1pt]b24.north east) -- ([xshift=1pt]b17.north west)
node [pos=0.5,anchor=south,yshift=-2pt] {\footnotesize third byte};

\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=2pt
    },
    decorate
] ([xshift=1pt]b2.south west) -- ([xshift=-1pt]b4.south east)
node [pos=0.5,yshift=-2pt,anchor=north] (type) {\footnotesize object type};

\draw[<-] (b1) -- (type.south -| b1) node[anchor=north,align=center,font=\footnotesize] {if the MSB is one,\\then this is not\\the last byte};

\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=2pt
    },
    decorate
] ([xshift=1pt]b5.south west) -- ([xshift=-1pt]b8.south east)
node [pos=0.5,yshift=-2pt,anchor=north,align=center,font=\footnotesize] (length) {first four\\bits of\\the length\\(big-endian)};

\draw[<-] (b9) -- (length.south -| b9) node[anchor=north,align=center,font=\footnotesize] {MSB is one,\\so this is \emph{not} the last byte};

\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=2pt
    },
    decorate
] ([xshift=1pt]b10.south west) -- ([xshift=-1pt]b16.south east)
node [pos=0.5,yshift=-2pt,anchor=north,align=center,font=\footnotesize] (next) {the next seven\\bits of the length\\(big-endian)};

\draw[<-] (b17) -- (next.south -| b17) node[anchor=north,align=center,font=\footnotesize] {MSB is zero,\\so this is the last byte};

\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=2pt
    },
    decorate
] ([xshift=1pt]b18.south west) -- ([xshift=-1pt]b24.south east)
node [pos=0.5,yshift=-2pt,anchor=north,align=center,font=\footnotesize] {the next seven\\bits of the length\\(big-endian)};

\end{tikzpicture}}
\end{center}
This example header represents an object of type \texttt{0b011}, which
happens to be a git blob, and an \emph{uncompressed} length of
$(100_2\,\verb|<<|\,14) +
(1010110_2\,\verb|<<|\,7) + 1001001_2
= 76617$ bytes.  Since this is not a delta object, it is immediately
followed by the zlib-compressed object data.

%% TODO: Talk about the Delta Ref file format

\section{A Minimal Polyglot PoC}

We now know that a git bundle is really just a git packfile with an
additional header, and a git packfile stores individual objects using
zlib, which uses the DEFLATE compression algorithm. DEFLATE supports
zero compression, so if we can store the PDF in a single object (as
opposed to it being split into deltas), then we could theoretically
coerce it to be intact within a valid git bundle.

Forcing the PDF into a single object is easy: We just need to add it
to the repo last, immediately before generating the bundle.

Getting the object to be compressed with zero compression is also
relatively easy.  That's because git was built in almost religious
adherence to The UNIX Philosophy: It is architected with hundreds of
sub commands it calls ``plumbing,'' of which the vast majority you
will likely have never heard. For example, you might be aware
that \texttt{git pull} is equivalent to a \texttt{git fetch} followed
by a \texttt{git merge}. In fact, the \texttt{pull} code is actually
spawning a new \texttt{git} child process to execute each of those
subcommands. Likewise, the \texttt{git bundle} command spawns
a \texttt{git pack-objects} to generate the packfile portion of the
bundle.  All we need to do is inject the \texttt{--compression=0}
argument into the list of command line arguments passed
to \texttt{pack-objects}.  This is a one-line addition
to \texttt{bundle.c}:\\
{\footnotesize
{\color{gray}
\verb|    argv_array_pushl(&pack_objects.args,|\\
\verb|                     "pack-objects", "--all-progress-implied",|}\\
\verb|                     "--compression=0",|\\
{\color{gray}
\verb|                     "--stdout", "--thin", "--delta-base-offset",|\\
\verb|                     NULL);|
}}

Using our patched version of git, every object stored in the bundle
will be uncompressed!
\begin{center}
\begingroup
\setbox9=\hbox{\footnotesize\verb|$ git bundle create article_bundle.pdf --all|}
\begin{minipage}{\wd9}
\footnotesize\begin{verbatim}
$ export PATH=/path/to/patched/git:$PATH
$ git init .
$ git add article.pdf
$ git commit article.pdf -m "added"
$ git bundle create polyglot.pdf --all
\end{verbatim}
\end{minipage}
\endgroup
\end{center}

\end{document}