### 2017-05-03 article.tex
Minor wording tweaks

\documentclass{article}

\pdfminorversion=4

\usepackage{tikz}
\usetikzlibrary{fit,patterns,decorations.pathreplacing}

\usepackage{graphicx}

\def\thindivider{\centerline{\tiny%%
-- --- --- -- ~~~ --- --- --- ~~~ --- -- --- -- ~~~~~~
--- --- --- ~~~  -- --- -- ~~~~~~
--- --- -- ~~~ --- ~~~ -- -- --- -- ~~~ --- --- ---}}

\begin{document}

\title{This PDF is a Git Repository\\Containing its Own \LaTeX\ Source\\and a Copy of Itself}
\author{Evan Sultanik}
\date{April 11, 2017}

\maketitle

Have you ever heard of the \texttt{git bundle} command? I hadn't. It
bundles a set of Git objects---potentially even an entire
repository---into a single file. Git allows you to treat that file as
if it were a standard Git database, so you can do things like clone a
repo directly from it. Its purpose is to easily sneakernet pushes or
even whole repositories across air gaps.

\thindivider

Neighbors, it's possible to create a PDF that is also a Git
repository.

\begin{center}
\begingroup
\setbox9=\hbox{\footnotesize\verb|Receiving objects: 100% (174/174), 103.48 KiB | 0 bytes/s, done.|}
\begin{minipage}{\wd9}
\footnotesize\begin{verbatim}
$ git clone article.pdf foo
Cloning into 'foo'...
Receiving objects: 100% (174/174), 103.48 KiB | 0 bytes/s, done.
Resolving deltas: 100% (100/100), done.
$ cd foo
$ ls
article.tex
article.pdf
\end{verbatim}
\end{minipage}
\endgroup
\end{center}

\section{The Git Bundle File Format}
\label{sec:GitBundleFormat}

The file format for Git bundles doesn't appear to be formally
specified anywhere, however, inspecting \texttt{bundle.c} reveals that
it's relatively straightforward:
\def\returnkey{\hspace*{1pt}{\setbox9=\hbox{$\hookleftarrow$}\tikz\node[text width=\wd9,fill=lightgray,rounded corners=1pt] at (0,0){$\hookleftarrow$};}\hspace*{1pt}}
\begin{center}
\begin{tikzpicture}
  \node[text width=0.9\hsize] at (0,0) (sig) {\footnotesize\tt\# v2 git bundle{\tiny\returnkey}};
  \node[inner sep=0mm,fit=(sig),draw,rounded corners,dashed,minimum width=0.9\hsize] (sigbox) {};
  \node[draw,rounded corners,fill=white] at (sigbox.north) (siglabel) {\footnotesize Git Bundle Signature};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (sig.south west) (b1) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/heads/master{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b1.south west) (b2) {\footnotesize\tt 3aa340a2e3d125ab6703e5c9bdfede2054a9c0c5 refs/remotes/origin/master{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b2.south west) (b3) {\footnotesize\tt 4146cfe2fe9249fc14623f832587efe197ef5d2d refs/stash{\tiny\returnkey}};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b3.south west) (b4) {\footnotesize\tt babdda4735ef164b7023be3545860d8b0bae250a HEAD{\tiny\returnkey}};
  \node[inner sep=0mm,fit=(b1) (b4),draw,rounded corners,dashed,minimum width=0.9\hsize] (digestbox) {};
  \node[rotate=-90,draw,rounded corners,fill=white] at (digestbox.east) {\footnotesize Digest};
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (b4.south west) (empty) {{\tiny\returnkey}\footnotesize\ };
  \node[text width=0.9\hsize,anchor=north west,yshift=-0.5mm] at (empty.south west) (pack) {{\footnotesize\tt PACK}$\ldots$};
  \node[inner sep=0mm,fit=(pack),draw,rounded corners,dashed,minimum width=0.9\hsize] (packbox) {};
  \node[draw,rounded corners,fill=white] at (pack.south) {\footnotesize Git Packfile};
\end{tikzpicture}
\end{center}
Git has another custom format called a \emph{Packfile} that it uses to
compress the objects in its database, as well as to reduce network
bandwidth when pushing and pulling. The packfile is therefore an
obvious choice for storing objects inside bundles. This of course
raises the question: What is the format for a Git Packfile?

Git does have some internal documentation in
\begin{center}
  \texttt{Documentation/technical/pack-format.txt}
\end{center}
however, it is rather sparse, and does not provide enough detail to
fully parse the format. The documentation also has some
``observations'' that suggest it wasn't even written by the file
format's creator and instead was written by a developer who was later
trying to make sense of the code.

Luckily, Aditya Mukerjee already had to reverse engineer the file
format for his GitGo clean-room implementation of Git, and he wrote an
excellent blog entry about
it\footnote{\texttt{https://codewords.recurse.com/issues/three/unpacking-git-packfiles}}.

\begin{center}
{\small
\def\char#1{\,`\texttt{#1}'\,}
\setbox9=\hbox{\char{K}}
\newdimen\bytewidth\bytewidth=\wd9
\def\byte#1{\hbox to \bytewidth{\hfil\texttt{#1}\hfil}}
\def\desc#1#2{\hbox to #1\bytewidth{\hfil #2\hfil}}
\def\underbrace#1{\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=2pt
    },
    decorate
] ([xshift=1pt]#1.base west) -- ([xshift=-1pt]#1.base east) 
node (#1label) [pos=0.5,anchor=north,yshift=-2pt]}
\begin{tikzpicture}
\node[inner sep=0pt,minimum width=4\bytewidth] at (0,0) (pack) {\char{P}\char{A}\char{C}\char{K}};
\node[inner sep=0pt,anchor=base west] at (pack.base east) (version) {\byte{00}\byte{00}\byte{00}\byte{02}};
\node[inner sep=0pt,anchor=base west] at (version.base east) (numobj) {\desc{4}{\# objects}};
\underbrace{pack}{\tiny magic};
\underbrace{version}{\tiny version};
\underbrace{numobj}{\tiny big-endian 4 byte int};
\node[inner sep=0pt,anchor=north west] at (packlabel.south -| pack.west) (chunks) {one data chunk for each object};
\node[inner sep=0pt,anchor=north west] at ([yshift=-0.5\baselineskip]chunks.south west) (sha1) {20-byte SHA-1 of all the previous data in the pack};
\end{tikzpicture}}
\end{center}

Although not entirely required to understand the polyglot, I think it
is useful to describe the git packfile format here, since it is not
well documented elsewhere. If that doesn't interest you, it's safe to
skip to the next section.  But if you do proceed, I hope you like
Soviet holes, dear neighbor, because chasing this rabbit might remind
you of \includegraphics{kolskaya}.

\begin{center}
\includegraphics[width=0.33\hsize]{RazvodityeKrolikov_small}
\end{center}

Right, the next step is to figure out the ``chunk'' format.  The chunk
header is variable length, and can be as small as one byte. It encodes
the object's type and its \emph{uncompressed} size. If the object is
a \textit{delta} (\textit{i.e.}, a diff, as opposed to a complete
object), the header is followed by either the SHA-1 hash of the base
object to which the delta should be applied, or a byte reference
within the packfile for the start of the base object. The remainder of
the chunk consists of the object data, zlib-compressed.

This is the format of the variable length chunk header:
\begin{center}
{\newcount\bitnum\bitnum=0
\begin{tikzpicture}
\node[coordinate] at (0,0) (b0) {};
\foreach \i in {1,0,1,1,0,1,0,0,1,1,0,1,0,1,1,0,0,1,0,0,1,0,0,1} {
  {\newcount\prevbit\prevbit=\bitnum
  \global\advance\bitnum by 1
  \node[anchor=west] at (b\the\prevbit.east) (b\the\bitnum) {\i};}
  \draw ([xshift=1pt]b\the\bitnum.south west) -- ([xshift=-1pt]b\the\bitnum.south east);
}
\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=-2pt
    },
    decorate
] ([xshift=-1pt]b8.north east) -- ([xshift=1pt]b1.north west)
node [pos=0.5,anchor=south,yshift=-2pt] {\footnotesize first byte};
\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=-2pt
    },
    decorate
] ([xshift=-1pt]b16.north east) -- ([xshift=1pt]b9.north west)
node [pos=0.5,anchor=south,yshift=-2pt] {\footnotesize second byte};
\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=-2pt
    },
    decorate
] ([xshift=-1pt]b24.north east) -- ([xshift=1pt]b17.north west)
node [pos=0.5,anchor=south,yshift=-2pt] {\footnotesize third byte};

\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=2pt
    },
    decorate
] ([xshift=1pt]b2.south west) -- ([xshift=-1pt]b4.south east)
node [pos=0.5,yshift=-2pt,anchor=north] (type) {\footnotesize object type};

\draw[<-] ([yshift=-2pt]b1.south) -- (type.south -| b1) node[anchor=north,align=center,font=\footnotesize] {if the MSB is one,\\then this is not\\the last byte};

\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=2pt
    },
    decorate
] ([xshift=1pt]b5.south west) -- ([xshift=-1pt]b8.south east)
node [pos=0.5,yshift=-2pt,anchor=north,align=center,font=\footnotesize] (length) {first four\\bits of\\the length\\(big-endian)};

\draw[<-] ([yshift=-2pt]b9.south) -- (length.south -| b9) node[anchor=north,align=center,font=\footnotesize] {MSB is one,\\so this is \emph{not} the last byte};

\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=2pt
    },
    decorate
] ([xshift=1pt]b10.south west) -- ([xshift=-1pt]b16.south east)
node [pos=0.5,yshift=-2pt,anchor=north,align=center,font=\footnotesize] (next) {the next seven\\bits of the length\\(big-endian)};

\draw[<-] ([yshift=-2pt]b17.south) -- (next.south -| b17) node[anchor=north,align=center,font=\footnotesize] {MSB is zero,\\so this \emph{is} the last byte};

\draw [
    thick,
    decoration={
        brace,
        mirror,
        raise=2pt
    },
    decorate
] ([xshift=1pt]b18.south west) -- ([xshift=-1pt]b24.south east)
node [pos=0.5,yshift=-2pt,anchor=north,align=center,font=\footnotesize] {the next seven\\bits of the length\\(big-endian)};

\end{tikzpicture}}
\end{center}
The second through fourth most significant bits of the first byte are
used to store the object type. The remainder of the bytes in the
header are of the same format as bytes two and three in this
example. This example header represents an object of type $11_2$,
which happens to be a git blob, and an \emph{uncompressed} length of
$(100_2\,\verb|<<|\,14) + (1010110_2\,\verb|<<|\,7) + 1001001_2 =
76$,617 bytes.  Since this is not a delta object, it is immediately
followed by the zlib-compressed object data. The header does not
encode the \emph{compressed} size of the object, since the DEFLATE
encoding can determine the end of the object as it is being
decompressed.

At this point, if you found The Life and Opinions of Tristram Shandy
to be boring or frustrating, then it's probably best to skip to the
next section, 'cause it's turtles all the way down.

{\fontfamily{jkpvos}\selectfont\begin{center}
\begin{tikzpicture}
\node at (0,0) (text) {\begin{minipage}{0.8\hsize}\small
To come at the exact weight of things= in the scientific steel-yard, the fulchrum, [Walter Shandy] would say, should be almost invisible, to avoid all friction from popular tenets=,---without this= the minutiae of philosophy, which should always= turn the balance, will have no weight at all---Knowledge, like matter, he would affirm, was= divisible in infinitum---that the grains= and scruples= were as= much a part of it, as= the gravitation of the whole world.
\end{minipage}};
\node[anchor=north east] at (text.north west) {\LARGE``};
\node[anchor=south west] at (text.south east) {\LARGE''};
\end{tikzpicture}
\end{center}}

%% TODO: Talk about the Delta Ref file format

\section{A Minimal Polyglot PoC}

We now know that a git bundle is really just a git packfile with an
additional header, and a git packfile stores individual objects using
zlib, which uses the DEFLATE compression algorithm. DEFLATE supports
zero compression, so if we can store the PDF in a single object (as
opposed to it being split into deltas), then we could theoretically
coerce it to be intact within a valid git bundle.

Forcing the PDF into a single object is easy: We just need to add it
to the repo last, immediately before generating the bundle.

Getting the object to be compressed with zero compression is also
relatively easy.  That's because git was built in almost religious
adherence to The UNIX Philosophy: It is architected with hundreds of
sub commands it calls ``plumbing,'' of which the vast majority you
will likely have never heard. For example, you might be aware
that \texttt{git pull} is equivalent to a \texttt{git fetch} followed
by a \texttt{git merge}. In fact, the \texttt{pull} code actually
spawns a new \texttt{git} child process to execute each of those
subcommands. Likewise, the \texttt{git bundle} command spawns
a \texttt{git pack-objects} child process to generate the packfile
portion of the bundle.  All we need to do is inject
the \texttt{--compression=0} argument into the list of command line
arguments passed to \texttt{pack-objects}.  This is a one-line
addition to \texttt{bundle.c}:\\ {\footnotesize {\color{gray}
\verb|    argv_array_pushl(&pack_objects.args,|\\
\verb|                     "pack-objects", "--all-progress-implied",|}\\
\verb|                     "--compression=0",|\\
{\color{gray}
\verb|                     "--stdout", "--thin", "--delta-base-offset",|\\
\verb|                     NULL);|
}}

Using our patched version of git, every object stored in the bundle
will be uncompressed!
\begin{center}
\begingroup
\setbox9=\hbox{\footnotesize\verb|$ git bundle create article_bundle.pdf --all|}
\begin{minipage}{\wd9}
\footnotesize\begin{verbatim}
$ export PATH=/path/to/patched/git:$PATH
$ git init
$ git add article.pdf
$ git commit article.pdf -m "added"
$ git bundle create polyglot.pdf --all
\end{verbatim}
\end{minipage}
\endgroup
\end{center}
Any vanilla, un-patched version of git will be able to clone a repo
from the bundle. It will also be a valid PDF, since virtually all
PDF readers ignore garbage bytes before and after the PDF.

\section{Generalizing the PoC}

There are, of course, several limitations to the minimal PoC given in
the previous section:
\begin{enumerate}

\item Adobe, being Adobe, will refuse to open the polyglot unless the PDF is version 1.4 or earlier.  I guess it doesn't like some element of the git bundle signature or digest if it's PDF~1.5. Why? Because Adobe, that's why.

\item Leaving the entire Git bundle uncompressed is wasteful if the repo contains other files; really, we only need the PDF to be uncompressed.

\item If the PDF is larger than 65,535 bytes---the maximum size of an uncompressed DEFLATE block---then git will inject 5-byte deflate block headers inside the PDF, likely corrupting it.

\item Adobe will also refuse to open the polyglot unless the PDF is near the beginning of the packfile\footnote{Requiring the PDF header to start near the beginning of a file is common for many, but not all, PDF viewers.}.

\end{enumerate}

The first limitation is easy to fix by instructing \LaTeX\ to produce
a version 1.4~PDF by adding \texttt{\textbackslash pdfminorversion=4}
to the document.

The second limitation is a simple matter of software engineering,
adding a command line argument to the \texttt{git bundle} command that
accepts the hash of the single file to leave uncompressed, and passing
that hash to \texttt{git pack-objects}. I have created a fork of git
with this feature, available here:
\begin{center}
  \texttt{https://github.com/ESultanik/git/tree/UncompressedPack}
\end{center}

As an aside, while fixing the second limitation I discovered that if a
file has multiple PDFs concatenated after one another
(\textit{i.e.},~a git bundle polyglot with multiple uncompressed PDFs
in the repo), then the behavior is viewer-dependent: Some viewers will
render the first PDF, while others will render the last.  That's a fun
way to generate a PDF that displays completely different content in,
say, macOS~Preview versus Adobe.

The third limitation is very tricky, and ultimately why this polyglot
was not used for the PDF of a digital issue of PoC$\|$GTFO. I've a
solution, but it will not work if the PDF contains any objects
(\textit{e.g.}, images) that are larger than 65,535 bytes. A universal
solution would be to break up the image into smaller ones and tile it
back together, but that is not feasible for a document the size of a
PoC$\|$GTFO issue.

DEFLATE headers for uncompressed blocks are very simple: The first
byte encodes whether the following block is the last in the file, the
next two bytes encode the block length, and the last two bytes are the
ones' complement of the length.  Therefore, to resolve this issue, all
we need to do is move all of the DEFLATE headers that zlib created to
different positions that won't corrupt the PDF, and update their
lengths accordingly.

Where can we put a 5-byte DEFLATE header such that it won't corrupt
the PDF? We could use our standard trick of putting it in a PDF object
stream that we've exploited countless times before to enable
PoC$\|$GTFO polyglots. The trouble with that is: Object streams are
fixed-length, so once the PDF is decompressed (\textit{i.e.},~when a
repo is cloned from the git bundle), then all of the 5-byte DEFLATE
headers will disappear and the object stream lengths would all be
incorrect. Instead, I chose to use PDF comments, which start at any
occurrence of the percent sign character~(\%) outside a string or
stream and continue until the first occurrence of a newline.  All of
the PDF viewers I tested don't seem to care if comments include
non-ASCII characters; they seem to simply scan for a
newline. Therefore, we can inject ``\texttt{\%\textbackslash n}''
between PDF objects and move the DEFLATE headers there. The only
caveat is that the DEFLATE header itself can't contain a newline
byte~(\texttt{0x0A}), otherwise the comment would be ended
prematurely.  We can resolve that by adding extra spaces to the end of
the comment, increasing the length of the following DEFLATE block and
thus increasing the length bytes in the DEFLATE header and avoiding
the \texttt{0x0A}. The only concession made with this approach is that
PDF Xref offsets in the deflated version of the PDF will be off by a
multiple of 5, due to the removed DEFLATE headers. Fortunately, most
PDF readers can gracefully handle incorrect Xref offsets (at the
expense of a slower loading time), and this will only affect the PDF
contained in the repository, \emph{not} the PDF polyglot.

As a final step, we need to update the SHA-1 sum at the end of the
packfile~(\textit{q.v.}~Section~\ref{sec:GitBundleFormat}), since we
moved the locations of the DEFLATE headers, thus affecting the hash.

At this point, we have all the tools necessary to create a generalized
PDF/Git Bundle polyglot for \emph{almost} any PDF and git
repository. The only remaining hurdle is that some viewers require
that the PDF occur as early in the packfile as possible. At first, I
considered applying another patch directly the git source code to make
the uncompressed object first in the packfile. This approach proved to
be very involved, in part due to git's UNIX design philosophy and
architecture of generic code reuse. We're already updating the
packfile's SHA-1 hash due to changing the DEFLATE headers, so instead
I decided to simply reorder the objects after-the-fact, subsequent to
the DEFLATE header fix but before we update the hash. The only
challenge is that moving objects in the packfile has the potential to
break offset delta objects, since they refer to their base objects via
a byte offset within the packfile.  Moving the PDF to the beginning
will break any offset delta objects that occur after the original
position of the PDF that refer to base objects that occur before the
original position of the PDF. That's why I had to dive deeper into the
rabbit hole of the packfile format to understand the delta object
headers (as you saw at the end of Section~\ref{sec:GitBundleFormat},
if you were brave enough to finish it), since I originally attempted
to rewrite the broken ones. Rewriting the broken offset delta objects
is the \emph{correct} solution, but, in the end, I discovered a much
simpler way.

\begin{center}
\begin{tikzpicture}
\node at (0,0) (text) {\begin{minipage}{0.8\hsize}\small
As a matter of fact, G-d just questioned my judgement. He said,
`Terry, are you worthy to be the man who makes the temple? If you are,
you must answer: Is this [dastardly], or is this divine intellect?'
\end{minipage}};
\node[anchor=north east] at (text.north west) {\LARGE``};
\node[anchor=south west] at (text.south east) {\LARGE''};
\node[anchor=north east] at (text.south east) (terry) {\setbox9=\hbox{\llap{---}Terry A.\ Davis, creator of TempleOS}\parbox{\wd9}{\raggedright\box9 self-proclaimed ``smartest programmer that's ever lived''}};
\end{tikzpicture}
\end{center}

In the previous section, recall that we created the minimal PoC by
patching the command line arguments to \texttt{pack-objects}.  One of
the command line arguments that is already passed by default
is \texttt{--delta-base-offset}. Running \texttt{git help
pack-objects} reveals the following:
\begin{center}
{\footnotesize\setbox9\hbox{\verb|A packed archive can express the base object of a delta as either a|}
\begin{minipage}{\wd9}
\begin{verbatim}
A packed archive can express the base object of a delta as either a
20-byte object name or as an offset in the stream, but ancient
versions of Git don't understand the latter. By default, git
pack-objects only uses the former format for better compatibility.
This option allows the command to use the latter format for
compactness. Depending on the average delta chain length, this
option typically shrinks the resulting packfile by 3-5 per-cent.
\end{verbatim}
\end{minipage}}
\end{center}
So all we need to do is \emph{remove} the \texttt{--delta-base-offset}
argument and git will not include any offset delta objects in the
pack!

\end{document}